# Insert chapter 2 title here

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

# Data analysis for IODS week 2

```{r Data input, echo=F}
# Setting working directory with knitr.
library(knitr)
library(GGally)
opts_knit$set(root.dir ="C:/Users/Riku_L/Documents/GitHub/IODS-project/data")

# Read data

students2014 <- read.csv("data/learning2014.csv")

```

The data has been read into the system and it has been collected 3.12.2014 -- 10.1.2015. The original results were from an international survey of Approaches to Learning. The following data includes variables as described in [here](https://www.slideshare.net/kimmovehkalahti/the-relationship-between-learning-approaches-and-students-achievements-in-an-introductory-statistics-course-in-finland). Variables answer to questions like how deeply the individual is trying to learn and how organized is his/her studying.

Next, I will briefly examine the data set's dimensions and structure.

```{r Dimensions and structure}

# First six rows of the data set
head(students2014)

# The dimensions
dim(students2014)

# Type of variables
str(students2014)

# Summaríes of the variables included in the data
summary(students2014)


```

So, from the above we can conclude that the data has 166 observations and each observation consists of 7 variables. The data consists of mainly numerical information on the respondents

* gender
* age
* global attitude toward statistics
* points in the final exam 
* measures for
    + Deep learning
    + How surfaced is the learning
    + Strategic learning

Of the respondents 110 identified themselves as female and 56 as male. Respondents were 17-55 years old and half of them are 21-27 years old. They got an average of 22.72 points on the final exam and were somewhat deep learners ($\mu$ = 3.68). 

*Miten summamuuttujat luotu???*

## Graphical overview

First, we will construct pairwise plots of each variable and draw bigger histograms to see the variables' distribution.

```{r Plots, echo = F, out.width="\\maxwidth", out.height="\\maxheight"}
par(mfrow=c(3,2))
for (i in 2:7) hist(students2014[,i], main = colnames(students2014)[i])
par(mfrow=c(1,1))

ggpairs(students2014[,-1])
```

Summaries were printed in the previous section, but here we clearly see how the distribution of ages is heavily skewed to the left whereas other variables have approximately symmetrical distributions. From the plot matrix, we can conclude that none of the variables correlate with each other much as the highest absolute correlation of 0.437 is with Points and Attitude which would be classified as weak positive correlation.

## Regression

Let's perform a linear regression:

```{r Regression, echo =F}

# Model itself
model <- lm(Points ~ Attitude + Age + Stra_adj, data = students2014)

# Summary
summary(model)
```

From the summary it is visible that only attitude raises -- or affects -- performance in the final exam on 5 per cent significance level. For every attitude point gained it is estimated that an average of 0.35 exam points is achieved. We also notice from the output that age does not have a significant effect on our model so we remove it from the model.

```{r Update model}
model <- update(model, formula. = .~. -Age)

summary(model)
```

*Using a summary of your fitted model, explain the relationship between the chosen explanatory variables and the target variable (interpret the model parameters). Explain and interpret the multiple R squared of the model.*

```{r Diagnostics, out.width="\\maxwidth"}
par(mfrow=c(2,2))
plot(model, which=c(1,2,4,5))
```

*Produce the following diagnostic plots: Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage. Explain the assumptions of the model and interpret the validity of those assumptions based on the diagnostic plots. *